---
title: "Data processing"
author: "AnnaClaire Marley"
date: "4/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE}
# load packages
library(tidyverse) #version 1.3.0
library(raster) #Version 3.3-13
library(sdmpredictors) #Version 0.2.8
library(sf) #version 0.9-4
library(maps) #Version 3.3.0
library(rnaturalearth) #Version 0.1.0
library(rnaturalearthdata) #Version 0.1.0
library(rgdal) #Version 1.5-16

```

This Rmarkdown:

- imports layers of interest
- reprojects, crops, and masks them to the area of interest 
- If shapefile, reclassifies to raster with 1 for fill and 0 for not then exports as raster
- exports them

```{r}
# set up path for saving files 
path <- "../data/Active_Data_Layers/"

```


Area of Interest bounding box
```{r aoi}
# aoi shapefile
aoi <- read_sf("../data/gulf_of_mexico_aoi/",
                layer = "gulf_of_mexico_aoi-polygon")

# assign projections 
crs_aoi <- st_crs(aoi)

#projection
proj <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
```


Gulf area
```{r}
# load gulf area
gulf_area <- read_sf(dsn = "../data/gulf_mexico_area/",
                     layer = "iho")

# reproject
gulf_area_reproject <- st_transform(gulf_area, proj)

plot(gulf_area_reproject$geometry)
map('world',fill=T,add=T,col='gray')

```


EEZ 
```{r}
# read in world eez
eez <- read_sf(dsn = "../data/World_EEZ_v11_20191118/",
               layer = "eez_v11")

# filter to us eez and reproject
us_eez <- eez %>% 
  filter(TERRITORY1 == "United States") %>% 
  st_transform(proj)

# crop to the gulf
eez_gulf <- st_crop(us_eez, aoi) 
plot(eez_gulf$geometry)

# create new polygon of the eez that only counts as gulf area 
eez_gulf <- st_intersection(gulf_area_reproject, eez_gulf)
plot(eez_gulf$geometry)

write_sf(eez_gulf, paste0(path, "eez_gulf.shp"))

```

# Suitability Layers

### HYCOM data

#### Sea surface temperature

```{r}
path_hycom <- "../data/HYCOM/Year_2011/" # change year depending on which year you want to dowload

# max sst
max_sst_hycom <- list.files(path_hycom, pattern = "av_sst_max*") # run this for each new year
max_sst_hy_stack <- raster::stack(paste0(path_hycom, max_sst_hycom))

# min sst
min_sst_hycom <- list.files(path_hycom, pattern = "av_sst_min*") # run this for each new year
min_sst_hy_stack <- raster::stack(paste0(path_hycom, min_sst_hycom))

```

##### Annual SST
```{r}
# average annual maximum sst
max_sst_hy_an <- mean(max_sst_hy_stack)
plot(max_sst_hy_an)

# process and plot
max_sst_hy_an_mask <- process_bo_data(max_sst_hy_an)
plot(max_sst_hy_an_mask)

# average annual minimum sst
min_sst_hy_an <- mean(min_sst_hy_stack)

# process and plot
min_sst_hy_an_mask <- process_bo_data(min_sst_hy_an)
plot(min_sst_hy_an_mask)

# write to file
writeRaster(max_sst_hy_an_mask, paste0(path, "max_sst_hy_an_mask_2011.tif"), overwrite = T)
writeRaster(min_sst_hy_an_mask, paste0(path, "min_sst_hy_an_mask_2011.tif"), overwrite = T)

```

```{r}
# combine multiple years to get mean 2010-2012
path_al <- "../data/Active_Data_Layers/"
  # max sst
  max_sst_hycom_yrs <- list.files(path_al, pattern = "max_sst_hy_an_mask*") # run this for each new year
  max_sst_hy_yrs_stack <- raster::stack(paste0(path_al, max_sst_hycom_yrs))
  max_sst_hy_10_12 <- mean(max_sst_hy_yrs_stack)
  # min sst
  min_sst_hycom_yrs <- list.files(path_al, pattern = "min_sst_hy_an_mask*") # run this for each new year
  min_sst_hy_yrs_stack <- raster::stack(paste0(path_al, min_sst_hycom_yrs))
  min_sst_hy_10_12 <- mean(min_sst_hy_yrs_stack)
  
# write to file
writeRaster(max_sst_hy_10_12, paste0(path, "max_sst_hy_10_12.tif"), overwrite = T)
writeRaster(min_sst_hy_10_12, paste0(path, "min_sst_hy_10_12.tif"), overwrite = T)
```


Create an Empty Raster Using min SST as a Template
```{r}
# Reclassification matrix for empty raster
rcl_matrix <- c(-Inf, Inf, 0)

# Reclassify the depth layer to make it an empty raster
empty_raster <- reclassify(min_sst_hy_an_mask, rcl= rcl_matrix)

```

#### Salinity

```{r}
path_hycom <- "../data/HYCOM/Year_2012/" # change year depending on which year you want to dowload

# max sst
max_sal_hycom <- list.files(path_hycom, pattern = "av_sal_max*") # run this for each new year
max_sal_hy_stack <- raster::stack(paste0(path_hycom, max_sal_hycom))

# min sst
min_sal_hycom <- list.files(path_hycom, pattern = "av_sal_min*") # run this for each new year
min_sal_hy_stack <- raster::stack(paste0(path_hycom, min_sal_hycom))

```


##### Annual salinity

```{r}
# annual average maximum salinity
max_sal_hy_an <- mean(max_sal_hy_stack)

# process and plot
max_sal_hy_an_mask <- process_bo_data(max_sal_hy_an)
plot(max_sal_hy_an_mask)

# annual average minimum salinity
min_sal_hy_an <- mean(min_sal_hy_stack)

# process and plot
min_sal_hy_an_mask <- process_bo_data(min_sal_hy_an)
plot(min_sal_hy_an_mask)

# write to file
writeRaster(max_sal_hy_an_mask, paste0(path, "max_sal_hy_an_mask_2012.tif"), overwrite = T)
writeRaster(min_sal_hy_an_mask, paste0(path, "min_sal_hy_an_mask_2012.tif"), overwrite = T)
```

```{r}
# combine multiple years to get mean 2010-2012
path_al <- "../data/Active_Data_Layers/"
  # max salinity
  max_sal_hycom_yrs <- list.files(path_al, pattern = "max_sal_hy_an_mask*") # run this for each new year
  max_sal_hy_yrs_stack <- raster::stack(paste0(path_al, max_sal_hycom_yrs))
  max_sal_hy_10_12 <- mean(max_sal_hy_yrs_stack)
  # min salinity
  min_sal_hycom_yrs <- list.files(path_al, pattern = "min_sal_hy_an_mask*") # run this for each new year
  min_sal_hy_yrs_stack <- raster::stack(paste0(path_al, min_sal_hycom_yrs))
  min_sal_hy_10_12 <- mean(min_sal_hy_yrs_stack)
  
# write to file
writeRaster(max_sal_hy_10_12, paste0(path, "max_sal_hy_10_12.tif"), overwrite = T)
writeRaster(min_sal_hy_10_12, paste0(path, "min_sal_hy_10_12.tif"), overwrite = T)
```


### GEBCO 

Depth Data
```{r}
# Get depth data of the gulf 
depth <- raster("../data/GEBCO_2019_21_Apr_2020_800d0264f91d/gebco_2019_n31.291637420654297_s17.777252197265625_w-100.55442810058594_e-78.51070404052734.tif")

# reproject
depth_proj <- projectRaster(depth, crs = proj)

# crop depth to EEZ
depth_crop <- crop(depth, aoi)

```

Resample

```{r}
# Resample to Bio-Oracle Cell Size (ATM Bio-Oracle Cell Size is the Standard) #Ask AC??
  # make resample crop
  max_sst_crop <- crop(projectRaster(max_sst, crs = proj), aoi)
  
  # resample
  depth_resample_bo <- resample(depth_crop, max_sst_crop, method='ngb', progress='text')

# mask depth to EEZ
depth_mask_bo <- mask(depth_resample, eez_gulf)

plot(depth_mask_bo)

# Write a Raster for Suitability Analysis
writeRaster(depth_mask_bo, paste0(path, "depth_mask_bo.tif"), overwrite = T)

### Resample to HYCOMM cell size
  # make resample crop
  max_sst_hy <- crop(projectRaster(min_sst_hy_an, crs = proj), aoi)
  
  # resample
  depth_resample_hy <- resample(depth_crop, max_sst_hy, method='ngb', progress='text')

# mask depth to EEZ
depth_mask_hy <- mask(depth_resample_hy, eez_gulf)

plot(depth_mask_hy)

# Write a Raster for Suitability Analysis
writeRaster(depth_mask_hy, paste0(path, "depth_mask_hy.tif"), overwrite = T)


```



# Exclusion Layers

## Natural Resources

Marine Protected Areas

```{r}
# read in MPAs
mpa <- read_sf("../data/MPAI2017", layer = "MPAI_2017")

# reproject
mpa_project <- mpa %>% 
  dplyr::select(Site_ID, Site_Name, Gov_Level, State) %>% 
  st_transform(proj) %>% 
  st_buffer(0)

# Crop
mpa_crop <- st_crop(mpa_project, aoi)

# Rasterizing Steps for MPA Layer
  # Create a Binary MPA Raster
  mpas_binary <- rasterize(mpa_crop, empty_raster, field = 0, background = 1) %>% 
    mask(eez_gulf)
  
  # Plot Reclassified MPA Layer
  plot(mpas_binary)
  #freq(mpas_binary)

# Write a Raster for Suitability Analysis
writeRaster(mpas_binary, paste0(path, "mpas_binary_hy.tif"), overwrite = T)
    
```

## Travel lanes

Shipping lanes
```{r}
# read in shipping lanes
shipping <- read_sf("../data/shippinglanes/", layer = "shippinglanes")

# reproject 
ship_project <- shipping %>% 
  st_transform(proj) 

# crop
ship_crop <- ship_project %>% 
  st_crop(eez_gulf)

# filter to only include shipping lanes and fairways
ship_filter <- ship_crop %>% 
  filter(THEMELAYER == "Shipping Fairways Lanes and Zones" | 
           THEMELAYER == "Traffic Separation Schemes" |
           THEMELAYER == "Traffic Separation Schemes/Traffic Lanes")

# rasterize & mask
ship_binary <- rasterize(ship_filter, empty_raster, field = 0, background = 1) %>% 
    mask(eez_gulf)

plot(ship_binary)

# Write a Raster for Suitability Analysis
writeRaster(ship_binary, paste0(path, "ship_binary_hy.tif"), overwrite = T)

```

## Infrastructure

Oil and gas platforms

```{r}
# read in oil and gas platforms 
og_platform <- read_sf(dsn = "../data/OilandGasPlatforms/OilandGasPlatforms.gdb/",
                       layer = "OilandGasPlatforms")

# add 500 meter buffer around points
og_buffer <- st_buffer(og_platform, 500)

# reproject and crop
og_crop <- og_buffer %>% 
  st_transform(proj) %>% 
  st_crop(aoi)

# rasterize & mask
og_binary <- rasterize(og_crop, empty_raster,
                                   field = 0, background = 1) %>% 
  mask(eez_gulf)

# plot it
plot(og_binary)
map('world',fill=T,add=T,col='gray')

# Write a Raster for Suitability Analysis
writeRaster(og_binary, paste0(path, "og_binary_hy.tif"), overwrite = T)
```

Submarine cable areas

```{r}
# read in submarine cables
sub_cable_area <- read_sf(dsn = "../data/SubmarineCableAreas/SubmarineCableAreas.gdb/",
                      layer = "SubmarineCableAreas" )

# reproject 
sub_cable_area_project <- sub_cable_area %>% 
  st_transform(proj) 

# crop
sub_cable_area_crop <- st_crop(sub_cable_area_project, aoi)

plot(sub_cable_area_crop)

# rasterize & mask
sub_cable_area_binary <- rasterize(sub_cable_area_crop, empty_raster,
                                   field = 0, background = 1) %>% 
  mask(eez_gulf)

plot(sub_cable_area_binary)
map('world',fill=T,add=T,col='gray')

# Write a Raster for Suitability Analysis
writeRaster(sub_cable_area_binary, paste0(path, "sub_cable_area_binary_hy.tif"), overwrite = T)

```

Pipeline areas
```{r}
# read in pipeline areas
pipeline_areas <- read_sf(dsn = "../data/PipelineArea/PipelineArea.gdb/",
                          layer = "PipelineArea")

# reproject
pipe_project <- st_transform(pipeline_areas, proj)

# crop
pipe_crop <- st_crop(pipe_project, aoi)

plot(pipe_crop$Shape)

# rasterize and mask
pipe_binary <- rasterize(pipe_crop, empty_raster,
                                   field = 0, background = 1) %>% 
  mask(eez_gulf)

plot(pipe_binary)

# Write a Raster for Suitability Analysis
writeRaster(pipe_binary, paste0(path, "pipe_binary_hy.tif"), overwrite = T)

```

## Military

Danger and restricted zones
```{r}

# read in sf

danger_zones <- st_read(dsn = "../data/DangerZonesAndRestrictedAreas/DangerZonesandRestrictedAreas.gdb/", 
                        layer = "DangerZonesandRestrictedAreas")

# reproject
danger_reproject <- st_transform(danger_zones, proj) 

# recast to multipolygon for cropping
danger_mp <- st_cast(danger_reproject, "MULTIPOLYGON")

# crop
danger_crop <- st_crop(danger_mp, aoi)

# rasterize and mask
danger_binary <- rasterize(danger_crop, empty_raster,
                                   field = 0, background = 1) %>% 
  mask(eez_gulf)

plot(danger_binary)

# Write a Raster for Suitability Analysis
writeRaster(danger_binary, paste0(path, "danger_binary_hy.tif"), overwrite = T)

```

All suitable areas
```{r}

suitable <- overlay(ship_binary, mpas_binary, sub_cable_area_binary, pipe_binary,
                    danger_binary, og_binary, 
                    fun = function(a, b, c, d, e, f){a*b*c*d*e*f})
plot(suitable)
map('world',fill=T,add=T,col='gray')
```

## Need to figure out buffering -- ASK AC??

Submarine cables

```{r}
# read in submarine cables
sub_cables <- read_sf(dsn = "../data/SubmarineCables/NOAAChartedSubmarineCables.gdb/",
                      layer = "NOAAChartedSubmarineCables")

# reproject 
sub_cables_project <- sub_cables %>% 
  st_transform(proj) 

# recast
sub_cables_cast <- st_cast(sub_cables_project, "MULTILINESTRING")

# crop
sub_cables_crop <- st_crop(sub_cables_cast, aoi)

plot(sub_cables_crop$Shape)

# rasterize & mask
sub_cables_binary <- rasterize(sub_cables_crop, empty_raster,
                                   field = 0, background = 1) %>% 
  mask(eez_gulf)


plot(sub_cables_binary)

```

Oil and gas wells

```{r}
# read in oil and gas wells
#og_wells <- read_sf(dsn = "../data/OilandGasWells/OilandGasWells.gdb/",
 #                   layer = "OilandGasWells")

# reproject 
#og_wells_project <- og_wells %>% 
 # st_transform(crs_aoi)

```

Artificial Reefs
```{r}
# load artificial reefs from ocean reports
artificial_reefs <- sf::st_read(dsn = "../data/ArtificialReefs/ArtificialReefs.gdb", 
                                layer = "ArtificialReefs") %>% 
  st_transform(proj)

plot(artificial_reefs)

# Rasterize artificial reefs
#reefs_artificial_binary <- rasterize(artificial_reefs, empty_raster, field = 0, background = 1) %>%
 # mask(eez_BRA)


```





