---
title: "HYCOM Suitability"
author: "AnnaClaire Marley"
editor: "Phoebe Racine"
date: "6/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#load packages
library(tidyverse) #Version 1.3.0
library(raster) #Version 3.3.13
library(sdmpredictors) #Version 0.2.8
library(sf)  #Version 0.9.4
library(maps) #Version 3.3.0

```
# Suitability 

**General Seaweed:**
- Depth: 10-200m
- N:P ratio: between 4 and 80
-----------------------------------------
- Sea surface temperature range: 0-35°C
- Current velocity: 0.1-1 m/s

*Eucheuma* specific [1-3]:
- Optimal sea surface temp: 18-30°C
- Optimal salinity: 30-40 psu

*Gracilaria* specific [4,5]:
- Optimal sea surface temp: 15-34°C
- Optimal salinity: 8-60 psu
- 
*Sargassum* specific [6]:
- Optimal sea surface temp: 18-30°C
- Optimal salinity: 24-42 psu


**References**
1. Dawes, C. J. (1989). Temperature acclimation in cultured Eucheuma isiforme from Florida and E. alvarezii from the Philippines. Journal of Applied Phycology, 1(1), 59.

2. Tiensongrusmee, B. Site Selection for Eucheuma. Food and Agriculture Organization. http://www.fao.org/3/AB738E/AB738E04.htm

3. Dawes, C. J., Mathieson, A. C., & Cheney, D. P. (1974). Ecological studies of Floridian Eucheuma (Rhodophyta, Gigartinales). I. Seasonal growth and reproduction. Bulletin of Marine Science, 24(2), 235-273.

4. Bird, N. L., Chen, L. M., & McLachlan, J. (1979). Effects of temperature, light and salinity on growth in culture of Chondrus crispus. Furcellaria lumbricalis, Gracilaria tikvahiae (Gigartinales, Rhodophyta), and Fucus serratus (Fucales, Phaeophyta). Botanica Marina, 22(8), 521-528.

5. Gorman, L., Kraemer, G. P., Yarish, C., Boo, S. M., & Kim, J. K. (2017). The effects of temperature on the growth rate and nitrogen content of invasive Gracilaria vermiculophylla and native Gracilaria tikvahiae from Long Island Sound, USA. Algae, 32(1), 57-66.

6. Rocha, C. M., Sousa, A. M., Kim, J. K., Magalhães, J. M., Yarish, C., & do Pilar Gonçalves, M. (2019). Characterization of agar from Gracilaria tikvahiae cultivated for nutrient bioextraction in open water farms. Food Hydrocolloids, 89, 260-271.



**Seaweed Parameters**
```{r}
# depth
min_depth <- -10
max_depth <- -200

# Nitrate:Phosphate ratio
n_p_ratio_min <- 4
n_p_ratio_max <- 80

# sea surface temperature
min_sst_value <- 18
max_sst_value <- 30

# salinity
min_sal_value <- 30
max_sal_value <- 40

```

```{r}
# set up path to load activate layers
path <- "../data/Active_Data_Layers/"
```

```{r}
# load function to relcassify rasters
source('../functions/reclass_data.R')
```

```{r}
eez_gulf <- read_sf(path, layer = "eez_gulf")
```


Depth
```{r}

# read in raster
depth_mask <- raster(paste0(path,"/depth_mask_hy.tif"))

# Reclassification matrix 
rcl_matrix_depth <- c(-Inf, max_depth, 0,
                   max_depth, min_depth, 1,
                   min_depth, 0, 0,
                   0, Inf, 0)

# Reclassify 
depth_binary <- reclassify(depth_mask, rcl = rcl_matrix_depth)

# Plot reclassified layer
plot(depth_binary)

```


Max SST
```{r}
# read in raster
max_sst <- raster(paste0(path,"/max_sst_hy_10_12.tif"))


# reclassify
max_sst_binary <- reclass_data(raster = max_sst, max_thresh = max_sst_value, 
                        one_thresh = TRUE, which_thresh = "MAX")


# plot reclassified layer
plot(max_sst_binary)

```


Min SST
```{r}
# read in raster
min_sst <- raster(paste0(path,"/min_sst_hy_10_12.tif"))

plot(min_sst)

# reclassify
min_sst_binary <- reclass_data(raster = min_sst, min_thresh = min_sst_value, 
                        one_thresh = TRUE, which_thresh = "MIN")


# plot reclassified layer
plot(min_sst_binary)
summary(min_sst)

```

Salinity
```{r}
# read in files
sal_max <- raster(paste0(path,"/max_sal_hy_10_12.tif"))
sal_min <- raster(paste0(path,"/min_sal_hy_10_12.tif"))

plot(sal_min)
plot(sal_max)

# reclassify
sal_max_binary <- reclass_data(raster = sal_max, max_thresh = max_sal_value, 
                        one_thresh = TRUE, which_thresh = "MAX")
sal_min_binary <- reclass_data(raster = sal_min, min_thresh = min_sal_value, 
                        one_thresh = TRUE, which_thresh = "MIN")

# plot
plot(sal_max_binary)
plot(sal_min_binary)

```

```{r,echo = FALSE}
plot(depth_binary)
plot(max_sst_binary)
plot(min_sst_binary)
plot(sal_max_binary)
plot(sal_min_binary)
```




N:P ratio
```{r}
# read in raster
n_p_ratio <- raster(paste0(path,"/n_p_ratio.tif"))

# resample
n_p_ratio_resample_hy <- resample(n_p_ratio, sal_max, method='ngb', progress='text')

# reclassify
np_binary <- reclass_data(raster = n_p_ratio_resample_hy, min_thresh = n_p_ratio_min,
                          max_thresh = n_p_ratio_max, one_thresh = FALSE)

# plot reclassified layer
plot(np_binary)
```





### Final Maps

```{r}
# source functions to save maps created
source("../functions/save_map.R")
```

Suitability areas -- all bio-oracle
```{r}

suitable_areas <- raster::overlay(np_binary,
                                  depth_binary, 
                                  max_sst_binary,
                                  min_sst_binary, 
                                  sal_max_binary, 
                                  sal_min_binary,
                          fun = function(a, b, c, d, e, f){a*b*c*d*e*f})

suitable_areas <- raster::overlay(
                                  depth_binary, 
                                  max_sst_binary,
                                  min_sst_binary, 
                                  sal_max_binary, 
                                  sal_min_binary,
                          fun = function(a, b, c, d, e){a*b*c*d*e})


suit_area <- tapply(area(suitable_areas), suitable_areas[], sum)

# zoom in to see
suitable_zoom <- crop(suitable_areas, extent(-100, -80, 22, 35))


# plot
plot(suitable_zoom, legend = FALSE, col = c("white", "paleturquoise4"))
plot(eez_gulf$geometry, add = T, lty=2)
map('world',fill=T,add=T,col='gray')
map('state',fill=F,add=T)


```

```{r}
### save the map ###
#save_map(png_name = "seaweed_gen_suitable.png", title = "General Seaweed Suitability", 
#         color = "light blue")

# Eucheuma
#save_map(png_name = "eucheuma_suitable.png", title = "Eucheuma Suitability", 
#         color = "green")
# Sargassum
#save_map(png_name = "sargassum_suitable_v6.png", title = "Sargassum Suitability", 
 #        color = "dark green")
# Gracilaria
#save_map(png_name = "gracilaria_suitable_v2.png", title = "Gracilaria Suitability", 
 #        color = "purple")
```

# Exclusion Layers

```{r}
# Marine protected areas
mpa <- raster(paste0(path,"/mpas_binary_hy.tif"))

# Shipping lanes
shipping_lanes <- raster(paste0(path,"/ship_binary_hy.tif"))

# Submarine Cable Areas
sub_cable_areas <- raster(paste0(path,"/sub_cable_area_binary_hy.tif"))

# Pipeline Areas
pipeline_areas <- raster(paste0(path,"/pipe_binary_hy.tif"))

# Military danger zones
military_danger <- raster(paste0(path,"/danger_binary_hy.tif"))

# Oil and Gas platforms
oil_gas_platforms <- raster(paste0(path,"/og_binary_hy.tif"))

```

# Final Suitability Map
```{r}

exclusion_areas <- overlay(
                          mpa,
                          shipping_lanes,
                          sub_cable_areas,
                          pipeline_areas,
                          military_danger,
                          oil_gas_platforms,
                          fun = function(a, b, c, d, e, f){a*b*c*d*e*f})

suitable_areas <- overlay(np_binary, 
                          depth_binary, 
                          max_sst_binary,
                          min_sst_binary, 
                          sal_max_binary, 
                          sal_min_binary,
                          mpa,
                          shipping_lanes,
                          sub_cable_areas,
                          pipeline_areas,
                          military_danger,
                          oil_gas_platforms,
                          fun = function(a, b, c, d, e, f, g, h, i, j,  k, l){a*b*c*d*e*f*g*h*i*j*k*l})

plot(suitable_areas)
map('world',fill=T,add=T,col='gray')
map('state',fill=F,add=T)

suit_exclus_area <- tapply(area(suitable_areas), suitable_areas[], sum)

# zoom in to see
exclusion_zoom <- crop(suitable_areas, extent(-100, -80, 22, 35))

plot(exclusion_zoom, legend = FALSE, col = c("white", "paleturquoise4"))
plot(eez_gulf$geometry, add = T, lty=2)
map('world',fill=T,add=T,col='gray')
map('state',fill=F,add=T)

```
